\cleardoublepage
\chapter*{Notations}
\markboth{Notations}{Notations}
\addcontentsline{toc}{chapter}{Notations}

\begin{itemize}
    \item $\N$ denotes the set of natural integers and $\R$ the set of real numbers.
    \item Vectors are written in bold letters, e.g. $\xx,\, \yy \in \R^p$.
    \item Matrices are written in capital letters, e.g. $X \in \R^{n \times p}$.
    \item We may cast any vector $\xx \in \R^p$ to a column matrix $x \in \R^{p \times 1}$ with the same entry values.
    \item We note $\odot$ the Hadamard (element-wise) product between two vectors or matrices.
    \item $\0$ and $\1$ are the vectors whose all entries are $0$ and $1$ respectively, and whose size is inferred by the context.
    \item For a vector $\vv$, we note $\norm{\vv}_0$ the number of non-zero entries of $\vv$ (that is, its cardinality).
    \item The $\diag$ operator can be used in two contexts;
    either to transform a vector into a diagonal matrix whose diagonal entries match the vector,
    or to extract the diagonal of a matrix.
    For example
    \begin{equation*}
        \diag \begin{bmatrix}
                  1\\
                  2
        \end{bmatrix}
        =
        \begin{bmatrix}
            1 & 0\\
            0 & 2
        \end{bmatrix}
        ,\qquad
        \diag \begin{bmatrix}
                  1 & 2\\
                  3 & 4
        \end{bmatrix}
        =
        \begin{bmatrix}
            1\\
            4
        \end{bmatrix}
    \end{equation*}
    \item Given two matrices $A$ and $B$, we note $\big[ A, B \big]$ and $\big[ A; B \big]$ their horizontal and
    vertical concatenation respectively (if dimensions match).
    \item Functions $\R \to \R$ applied to matrices are implicitly performed element-wise, unless otherwise specified.
    \item $\Eb{\cX}$ denotes the expectation of a random variable (or vector) $\cX$.
    \item Given a set $\cS$, $\abs{\cS}$ and $\#\cS$ denote its cardinality.
    \item The sets of $n \times n$ symmetric matrices,
    symmetric positive semidefinite,
    and symmetric positive definite
    are denoted by $S^n$, $S^n_+$ and $S^n_{++}$ respectively.
\end{itemize}
